Hyperbolicity of Augmented Links in the Thickened Torus

### 2020-10-05 draft1.tex
made minor grammatical corrections

\begin{abstract}
For a hyperbolic link $K$ in the thickened torus, there is a decomposition of the complement
of a link $L$, obtained from augmenting $K$, into torihedra. Then we further decompose 
the torihedra into angled pyramids and finally angled tetrahedra. These fit into an angled structure 
on a triangulation of the link complement, and thus by \cite{Casson-Rivin}, this shows
that $L$ is hyperbolic.  
\end{abstract}

@@

Adams showed in \cite{CA} that given a hyperbolic alternating link $K$ in
$\Sp^3$ the link $L$ obtained by augmentation $K$ is hyperbolic. In this paper
we investigate if this statement holds for links in the thickened torus i.e. if
$L$ is a link obtained from augmenting a hyperbolic alternating link $K$ in the
thickened torus. We define augmenting similarly for links in the thickened torus 
with their associated link diagram on $\torus \times \{0\}.$

Menasco \cite{Menasco} showed there are decompositions of the complements of 
alternating links in $\Sp^3$ into two polyhedra, a top polyhedron and a bottom polyhedron. 
For alternating links, $K$ in the thickened torus Champanerkar, Kofman and Purcell \cite{CKP2}
showed that there is a decomposition of the complement of $K$ into objects called torihedra,
which we think of 
as counterparts of Menasco's decomposition of links in $\Sp^3$ into polyhedra, for links in the thickened 
torus. Just like Menasco's decomposition, one can obtain a top and bottom torihedron.
In \secref{s:auglinks} we show that for augmented links in the thickened torus 
one can also obtain a decomposition of the complement into a top and bottom torihedron.
In \secref{s:hyperbolicity}, we prove that augmented alternating links in the thickened torus
are hyperbolic.
In the Appendix, we provide an alternative proof to hyperbolicity of these links
via circle patterns, which one could use obtain lower bounds for volume.

@@

Now we can define augmentation for a link in $\torus \times I$ the same way we define augmentation for links in $\Sp^3$. For a link in $\torus \times I$, the crossing circles are added to the diagram projected onto $\torus \times \{0\}$. Let $L$ be a twist reduced diagram in $\torus \times I$, we define {\it augmentation} as a process in which an unknotted circle component is added to one or more twist regions of $L$. See Figure \ref{fig:Augmentations}

@@

We show a method of decomposing an augmented link in the thickened torus into
two torihedra. The idea is to combine methods of Menasco
\cite{Menasco} and the use of crossing edges between each crossing of our link
and Lackenby's ``cut-slice-flatten" method \cite{lackenby} on the augmentation
sites.

@@

\begin{proof}
We will begin by assuming that there are no half twists and then
arrange the link diagram of $L$ in the following way: first place the added
circle components (augmentation) perpendicular to the projection plane, $\torus
\times \{0\}$ leaving the remaining part of the link parallel to the projection
plane.  We now place a crossing edge on each crossing of the link so that for
each crossing edge, one end of the edge lies on a bottom strand while the other
end lies on a top strand as in Figure \ref{fig:crossingArc} left.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The following figures is an example which decomposes the link (C) of Figure
\ref{fig:Augmentations}.

\begin{figure}[h] 
\centering
\includegraphics[height=3cm]{fig-3}
\includegraphics[height=3cm]{fig-4}
	\caption{Each crossing circle bounds a disk}
\end{figure}

%\vspace{-0.5cm} 
\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-5} 
	\caption{We split the disk and collapse the arc of each
 crossing circle to ideal vertices} 
\end{figure}

\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-6} 
\caption{Left: The crossing arc is the edge in red.
Middle: Picture of splitting the crossing edge. Right: The link component is
pushed off to infinity.} 
\end{figure}

@@

Suppose the edges are labeled $e_i$,
which goes between vertices $v_i$ and $v_{i+1}$,
and suppose $e_i$ is assigned dihedral angle $\alpha_i$.
Let $e'$ be a new edge addeed to the base face of $P_n$
such that it separates the base face into a triangle and
an $(n-1)$-gon;
suppose the sides of the triangle are
$e_i, e_{i+1}$, and $e'$.
The new face corresponding to $e'$ separates $P_n$ into
an ideal tetrahedron $T$ and an ideal pyramid $P_{n-1}$.
We assign the dihedral angle of $\pi - \alpha_i - \alpha_{i+1}$
to $e'$ in $T$, and assign $\alpha_i + \alpha_{i+1}$ to $e'$ in $P_{n-1}$.
Clearly the sum of dihedral angles condition is satisfied
in $T$ and $P_{n-1}$.
It remains to check that the dihedral angles assigned to the vertical edges
are correct.
For the vertical edge associated to $v_j$ for $j \neq i, i+2$,
there is nothing to check;
for $j = i$, the dihedral angles are
$\pi - \alpha_i - (\pi - \alpha_i - \alpha_{i-1})$
in $T$ and $\pi - \alpha_{i-1} - (\alpha_i + \alpha_{i+1})$ in $P_{n-1}$,
which sum to $\pi - \alpha_i - \alpha_{i+1}$;
it is similar for $j = i+2$.

@@

For a hyperbolic link $K$ in $\torus \times I$, we show
that the resulting link obtained from augmenting $K$ is
hyperbolic. The idea is to start with a graph from the torihedral decomposition
of the link $K$ which will give us a graph on each torihedron with $\pi/2$ edges
\cite{CKP2},
the angled torihedra can be further decomposed into base-angled pyramids.
By \prpref{p:tori_decomp},
there is a torihedral decomposition of the complement of the augmented link $L$.
Using those angles,
we then assign angles to edges of these torihedra
and decompose them into base-angled pyramids.
These pyramids can be further decomposed into tetrahedra,
thus obtaining an angle structure on a triangulation.
In the Appendix, we show that this triangulation
also has an angle structure that
satisfies Thurston's gluing equations.

@@

We must check that this process terminates with a graph
which satisfies the hypotheses set out in the lemma.
It is clear that %checkboard colorability and
4-valency is maintained
at each step, and there will not be any bigons at the end.
We need to ensure that eliminating bigons does not
create self-loop edges.
Here we sepcify how to choose which way to resolve the internal
vertex $v$. If among the four faces touching $v$,
one of them has greater than three sides (Case 1),
then resolve $v$ so that this face is not merged;
if all four faces are triangles (Case 2),
simply resolve in an arbitrary manner.
After the modification of the internal vertex,
at most two bigons are created;
eliminate those two bigons.
Now we find that all bigons will be connected end-to-end
(i.e. forming a single twist region) -
this is because in Case 1, only one bigon was present after smoothing
$v$, which would result in at most two bigons that would be connected;
and if we were in Case 2, we find that the face between those
eliminated bigons (coming from merging the two triangles across $v$)
will become a bigon too,
thus connecting the potential bigons on either ends.

Thus, one can eliminate bigons in an order so that
they belong in one twist region;
in particular, the (not necessarily distinct) faces on either side
are not bigons,
so eliminating the bigon does not produce a self-loop edge.

@@

We obtain an angle-splitting by applying the
Feasible Flow theorem (\thmref{t:feasible_flow}) as follows:
Consider the directed graph whose vertex set is
$E \cup F \cup \{\otimes\}$,
where $E, F$ are the set of edges, faces in $\Gamma_T(K)$,
and $\otimes$ is some abstract vertex.
There is a directed edge
\begin{itemize}
\item $\otimes \to f$ for each face $f\in F$, with capacity interval
	$[\pi, \infty)$,
\item $f \to e$ for each edge $e \in \del f$,
	with capacity interval $[\veps, \infty)$
	for some $\veps>0$ to be set later,
\item $e \to \otimes$ for each edge $e$, with capacity interval
	$(-\infty, \pi/2]$.
\end{itemize}
By \lemref{l:2fe}, $2|F'| = 2\chi(F') \leq |E'|$,
and taking $\veps < \pi / |\text{max face size}|$,
the feasible flow condition is satisfied,
so a feasible flow exists.
Since $2|F| = |E|$, the capacity interval restrictions
on the flow at $\otimes$ is sharp, so out-edges at $\otimes$
have flow $\pi$ and in-edges at $\otimes$ have flow $\pi/2$.
Then the flow $f \to e$ gives us $\vphi_{\vec{e}}$,
where $f$ is the face to the left of $\vec{e}$.
(we adapted this argument from \cite{BandS}).

@@

\begin{remark}
\label{r:nohalftwist}
If the original link $K$ had some twist regions with at least one bigon,
we may consider augmentations $L$ where all such twist regions
are augmented.
Then, as pointed out in proof of \thmref{t:auglink_hyp} above,
the proof still works for $L$, showing that $L$ is also hyperbolic.
\end{remark}

>>>>

\begin{abstract}
For a hyperbolic link $K$ in the thickened torus, We show there is a decomposition of the complement
of a link $L$, obtained from augmenting $K$, into torihedra. We further decompose 
the torihedra into angled pyramids and finally angled tetrahedra. These fit into an angled structure 
on a triangulation of the link complement, and thus by \cite{Casson-Rivin}, this shows
that $L$ is hyperbolic.  
\end{abstract}

@@

Adams showed in \cite{CA} that given a hyperbolic alternating link $K$ in
$\Sp^3$ the link $L$ obtained by augmentating $K$ is hyperbolic. In this paper
we investigate if this statement holds for links in the thickened torus i.e. if
$L$ is a link obtained from augmenting a hyperbolic alternating link $K$ in the
thickened torus. We define augmenting similarly for links in the thickened torus 
with their associated link diagram on $\torus \times \{0\}.$

Menasco \cite{Menasco} showed there are decompositions of the complements of 
alternating links in $\Sp^3$ into two polyhedra, a top polyhedron and a bottom polyhedron. 
For alternating links, $K$ in the thickened torus Champanerkar, Kofman and Purcell \cite{CKP2}
showed that there is a decomposition of the complement of $K$ into objects called torihedra,
which we think of 
as counterparts to Menasco's decomposition of links in $\Sp^3$ into polyhedra, for links in the thickened 
torus. Just like Menasco's decomposition, one can obtain a top and a bottom torihedron.
In \secref{s:auglinks} we show that for augmented links in the thickened torus 
one can also obtain a decomposition of the complement into a top and bottom torihedron.
In \secref{s:hyperbolicity}, we prove that augmented alternating links in the thickened torus
are hyperbolic.
In the Appendix, we provide an alternative proof to hyperbolicity of these links
via circle patterns, which one could use to obtain lower bounds for volume.

@@

Now we can define augmentation for a link in $\torus \times I$ the same way we define augmentation for links in $\Sp^3$. For a link in $\torus \times I$, the crossing circles are added to the diagram projected onto $\torus \times \{0\}$. Let $L$ be a twist reduced diagram in $\torus \times I$, we define {\it augmentating} as a process in which an unknotted circle component is added to one or more twist regions of $L$. See Figure \ref{fig:Augmentations}

@@

We show a method of decomposing an augmented link in the thickened torus into
two objects called torihedra. The idea is to combine methods of Menasco
\cite{Menasco} and the use of crossing edges between each crossing of our link
and Lackenby's ``cut-slice-flatten" method \cite{lackenby} on the augmentation
sites.

@@

\begin{proof}
We will begin by assuming that there are no half twists. 
Arrange the link diagram of $L$ in the following way: first place the added
circle components perpendicular to the projection plane, $\torus
\times \{0\}$ leaving the remaining part of the link parallel to the projection
plane.  We now place a crossing edge on each crossing of the link so that for
each crossing edge, one end of the edge lies on a bottom strand while the other
end lies on a top strand as in Figure \ref{fig:crossingArc} left.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Figures \ref{fig:step_one} to \ref{fig:top-bottom} is an example which decomposes the link (C) of Figure
\ref{fig:Augmentations}.

\begin{figure}[h] 
\centering
\includegraphics[height=3cm]{fig-3}
\includegraphics[height=3cm]{fig-4}
	\caption{Each crossing circle bounds a disk}
	\label{fig:step_one}
\end{figure}

%\vspace{-0.5cm} 
\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-5} 
	\caption{We split the disk and collapse the arc of each
 crossing circle to ideal vertices}
	\label{fig:step_two}
\end{figure}

\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-6} 
\caption{Left: The crossing arc is the edge in red.
Middle: Picture of splitting the crossing edge. Right: The link component is
pushed off to infinity.}
	\label{fig:step_three}
\end{figure}

@@

Suppose the edges are labeled $e_i$,
which goes between vertices $v_i$ and $v_{i+1}$,
and suppose $e_i$ is assigned dihedral angle $\alpha_i$.
Let $e'$ be a new edge addeed to the base face of $P_n$
such that it separates the base face into a triangle and
an $(n-1)$-gon;
suppose the sides of the triangle are
$e_i, e_{i+1}$, and $e'$.
The new face corresponding to $e'$ separates $P_n$ into
an ideal tetrahedron $T$ and an ideal pyramid $P_{n-1}$.
We assign the dihedral angle of $\pi - \alpha_i - \alpha_{i+1}$
to $e'$ in $T$, and assign $\alpha_i + \alpha_{i+1}$ to $e'$ in $P_{n-1}$.
Clearly the sum of dihedral angles condition is satisfied
in $T$ and $P_{n-1}$.
	It remains to check that the dihedral angles assigned to the vertical (non-base)
	edges are correct.
For the vertical edge associated to $v_j$ for $j \neq i, i+2$,
there is nothing to check;
for $j = i$, the dihedral angles are
$\pi - \alpha_i - (\pi - \alpha_i - \alpha_{i-1})$
in $T$ and $\pi - \alpha_{i-1} - (\alpha_i + \alpha_{i+1})$ in $P_{n-1}$,
which sum to $\pi - \alpha_i - \alpha_{i+1}$;
it is similar for $j = i+2$.

@@

For a hyperbolic link $K$ in $\torus \times I$, we show
that the resulting link obtained from augmenting $K$ is
hyperbolic. The idea is to start with a graph from the torihedral decomposition
of the link $K$ which will give us a graph on each torihedron with an angle
assignment of
$\pi/2$ for edge
\cite{CKP2},
the angled torihedra can be further decomposed into base-angled pyramids.
By \prpref{p:tori_decomp},
there is a torihedral decomposition of the complement of the augmented link $L$.
Using those angles,
we then assign angles to edges of these torihedra
and decompose them into base-angled pyramids.
These pyramids can be further decomposed into tetrahedra,
thus obtaining an angle structure on a triangulation.
In the Appendix, we show that this triangulation
also has an angle structure that
satisfies Thurston's gluing equations.

@@

We must check that this process terminates with a graph
which satisfies the hypotheses set out in the lemma.
It is clear that %checkboard colorability and
4-valency is maintained
at each step, and there will not be any bigons at the end.
We need to ensure that eliminating bigons does not
create self-loop edges.
Here we specify how to choose which way to resolve the internal
vertex $v$. If among the four faces touching $v$,
one of them has greater than three sides (Case 1),
then resolve $v$ so that this face is not merged;
if all four faces are triangles (Case 2),
simply resolve in an arbitrary manner.
After the modification of the internal vertex,
at most two bigons are created;
eliminate those two bigons.
Now we find that all bigons will be connected end-to-end
(i.e. forming a single twist region) -
this is because in Case 1, only one bigon was present after smoothing
$v$, which would result in at most two bigons that would be connected;
and if we were in Case 2, we find that the face between those
eliminated bigons (coming from merging the two triangles across $v$)
will become a bigon too,
thus connecting the potential bigons on either ends.

Therefore, one can eliminate bigons in an order so that
they belong in one twist region;
in particular, the (not necessarily distinct) faces on either side
are not bigons,
so eliminating the bigon does not produce a self-loop edge.

@@

We obtain an angle-splitting by applying the
Feasible Flow theorem (\thmref{t:feasible_flow}) as follows:
Consider the directed graph whose vertex set is
$E \cup F \cup \{\otimes\}$,
where $E, F$ are the set of edges, faces of $\Gamma_T(K)$ respectively,
and $\otimes$ is some abstract vertex.
There is a directed edge
\begin{itemize}
\item $\otimes \to f$ for each face $f\in F$, with capacity interval
	$[\pi, \infty)$,
\item $f \to e$ for each edge $e \in \del f$,
	with capacity interval $[\veps, \infty)$
	for some $\veps>0$ to be set later,
\item $e \to \otimes$ for each edge $e$, with capacity interval
	$(-\infty, \pi/2]$.
\end{itemize}
By \lemref{l:2fe}, $2|F'| = 2\chi(F') \leq |E'|$,
and taking $\veps < \pi / |\text{max face size}|$,
the feasible flow condition is satisfied,
so a feasible flow exists.
Since $2|F| = |E|$, the capacity interval restrictions
on the flow at $\otimes$ is sharp, so out-edges at $\otimes$
have flow $\pi$ and in-edges at $\otimes$ have flow $\pi/2$.
Then the flow $f \to e$ gives us $\vphi_{\vec{e}}$,
where $f$ is the face to the left of $\vec{e}$.
(we adapted this argument from \cite{BandS}).

@@

\begin{remark}
\label{r:nohalftwist}
If the original link $K$ had some twist regions with at least one bigon,
we may consider augmentations $L$ where all such twist regions
are augmented, i.e. $L$ may have augmentations without half-twists.
Then, as pointed out in proof of \thmref{t:auglink_hyp} above,
the proof still works for $L$, showing that $L$ is also hyperbolic.
\end{remark}
